\documentclass{beamer}

\let\NOTARTICLE=1
\let\BEAMER=1
\let\FEWFONTS=1

\newcommand{\tick}{\checkmark}

\newenvironment{theindex}
 {\let\item\par
  %definitions for subitem etc
  }{}
\newcommand\indexspace{}

\input{../../common/preamble}

\newtheorem{proposition}{Proposition}

%%% END Article customizations

\title{Sequoidal Categories and Transfinite Games: A Coalgebraic Approach to Stateful Objects in Game Semantics}
\subtitle{University of Bath, 2017}

\author{John Gowers and James Laird}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}
  \frametitle{Semantics of Stateful Programs}
  In traditional domain-theoretic semantics, a function without side effects is represented as a map
  \[
    X \to Y
    \]
  \pause
  while a function with side effects is represented as a map
  \[
    S \times X \to S \times Y
    \]
  \pause
  or equivalently an algebra for the monad
  \[
    [S, S \times \blank]
    \]
  \pause
  We can then represent programs with a fixed store $S$ as morphisms in the Kleisli category for this monad.
  \pause
  This is an example of \emph{explicit state}.
\end{frame}

\begin{frame}
  \frametitle{Game Semantics and Implicit State}
  By contrast, the state in Game Semantics is \emph{implicit}.
  \pause

  Consider, for example, the game where player $O$, at any turn, can either say $b$ (button press) or $c$ (count).  
  
  Player $P$ must reply to $b$ with the move $\tick$ and can reply to $c$ by saying any natural number.
  \pause

  There is a strategy for this game corresponding to a simple counter.
  \pause

  An example play in that strategy is:
  \[
    c\ 0\ b\ \tick\ b\ \tick\ b\ \tick\ c\ 3\ c\ 3\ b\ \tick\cdots
    \]
  \pause
  This play has implicit state:
  \[
    \begin{array}{r|cccccccccccccc}
      \text{Moves} & c & 0 & b & \tick & b & \tick & b & \tick & c & 3 & c & 3 & b & \tick \\
      \text{State} & 0 & 0 & 1 & 1 & 2 & 2 & 3 & 3 & 3 & 3 & 3 & 3 & 3 & 3
    \end{array}
    \]
\end{frame}

\begin{frame}
  \frametitle{Case Study: The Abramsky-McCusker model of Idealized Algol}
  Abramsky and McCusker's model of Idealized Algol (1998) uses a stateful strategy $\cell$, representing the behaviour of a storage cell.
  \pause

  They use a combinatorial definition to define $\cell$ directly.  
  \pause

  Because state is implicit, it is not always easy to keep track of it.
  \pause

  Defining strategies combinatorially doesn't tell us what properties of the category of games make it suitable for modelling stateful programs.  
  \pause

  We want to find a way to \emph{encapsulate} an explicit state in a systematic category-theoretic way.
\end{frame}

\begin{frame}[fragile]
  \frametitle{Encapsulation of State}
  We can think of a state transformer as a box
  \[
    \begin{tikzpicture}
      \matrix[matrix of nodes, column sep=64pt, inner xsep=1ex, name=m, nodes={draw=black}] {
        & in & \\
        $S$ & $A$ & $S$ \\
        & out & \\};
      \path[commutative diagrams/.cd, every arrow, every label, labels={font size=tiny}]
        (m-2-1) edge node {old state} (m-2-2)
        (m-2-2) edge node {new state} (m-2-3);
    \end{tikzpicture}
    \]
  \pause
  Then \emph{encapsulation} means plugging an unbounded number of boxes together:
  \tikzset{dotsstyle/.style={draw=none}}
  \[
    \begin{tikzpicture}
      \matrix[matrix of nodes, column sep=36pt, inner xsep=1ex, name=m, nodes={draw=black}] {
        & in & in & in & & \\
        $S$ & $A_1$ & $A_2$ & $A_3$ & |[dotsstyle]| $\cdots$ \\
        & out & out & out & & \\};
      \path[commutative diagrams/.cd, every arrow, every label, labels={font size=tiny}]
        (m-2-1) edge node {init} (m-2-2)
        (m-2-2) edge node {state} (m-2-3)
        (m-2-3) edge node {state} (m-2-4)
        (m-2-4) edge node {state} (m-2-5);
    \end{tikzpicture}
    \]
\end{frame}

\begin{frame}
  \frametitle{Final coalgebras}
  A well-studied technique in semantics is to use a \emph{final coalgebra} in order to encapsulate state.  
  \pause

  If we have a functor corresponding to `adding  a box':
  \tikzset{bigbox/.style={minimum width=3cm, minimum height=2cm}}
  \[
    \begin{tikzpicture}
      \matrix[matrix of nodes, column sep=64pt, inner xsep=1ex, name=m, nodes={draw=black}] {
        & in & \\
        $S$ & $A$ & $S$ \\
        & out & \\};
      \path[commutative diagrams/.cd, every arrow, every label, labels={font size=tiny}]
        (m-2-1) edge node {old state} (m-2-2)
        (m-2-2) edge node {new state} (m-2-3);
    \end{tikzpicture}
    \]

\end{document}
