\documentclass{beamer}

\let\NOTARTICLE=1
\let\BEAMER=1
\let\FEWFONTS=1

\newenvironment{theindex}
 {\let\item\par
  %definitions for subitem etc
  }{}
\newcommand\indexspace{}

\input{../../common/preamble}

\newtheorem{proposition}{Proposition}

%%% END Article customizations

\title{Games for countable nondeterminism}
\subtitle{GaLoP 2017, Uppsala}

\author{John Gowers}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\section*{Outline}
\begin{frame}
  \tableofcontents
\end{frame}

\section{Game Semantics of Nondeterminism}

\renewcommand*{\arraystretch}{3}

\begin{frame}[allowframebreaks]
  \frametitle{Related Work}
  \nocite{mcCHFiniteND, RusssThesis, TsukadaSheaves, RoscoeCspInfinite, LevyGsInfinite, LairdOrdinalGames}
  \bibliographystyle{plain}
  \bibliography{../../common/phd_bibliography}
\end{frame}

\begin{frame}
  \frametitle{Nondeterministic Strategies}
  
  \pause
  If $A$ is a game with a set $P_A$ of positions, a strategy $\sigma$ for $A$ is a non-empty prefix-closed subset of $P_A^P$ satisfying:
  \begin{description}
    \item[Determinism] If $sa,sb\in\sigma$, then $a=b$
    \item[Totality] If $s\in\sigma$ and $sa\in P_A$, then $sab\in\sigma$ for some $b$
  \end{description}
  \pause
  Relaxing \emph{totality} gives us \emph{partial strategies}.
  \pause

  Relaxing \emph{determinism} gives us \emph{nondeterministic strategies}.
  \pause

  Problem: how can we model divergence in nondeterministic strategies?
\end{frame}

\begin{frame}
  \frametitle{The Harmer-McCusker model}
  \pause

  Harmer and McCusker model nondeterminism by equipping each (nondeterministic) strategy with an explicit set of divergent positions.  
  \pause

  Divergence in a composition of strategies may arise either from a divergence in one of the strategies we are composing, or through \emph{livelock} (infinite chattering).  
  \pause

  This method only works for finite nondeterminism: if we naively relax the finite branching condition, then composition is not associative.  
\end{frame}

\begin{frame}
  \frametitle{Failure of associativity in the presence of infinite branching}
  \begingroup
  \renewcommand*{\arraystretch}{1}
  \[
    \begin{array}{ccccccccc}
      \mathbf{I} & & \mathbf{\Sigma^*} & & \mathbf{\bN} & \tensor & \mathbf{\Sigma^*} & & \mathbf{\Sigma} \\
       & &   & &   & &   & & q \\
       & &   & & q & &   & &   \\
       & &   & & 2 & &   & &   \\
       & &   & &   & & q & &   \\
       & & q & &   & &   & &   \\
       & & a & &   & &   & &   \\
       & &   & &   & & a & &   \\
       & &   & &   & & q & &   \\
       & & q & &   & &   & &   \\
       & & a & &   & &   & &   \\
       & &   & &   & & a & &   \\
       & &   & &   & &   & & a
    \end{array}
    \]
  \endgroup
\end{frame}

\begin{frame}
  \frametitle{Solution: keep track of infinite sequences of moves}
  A game $A$ is now given by a tuple $(M_A, \lambda_A, \zeta_A, P_A)$ where
  \begin{itemize}
    \item $M_A$ is a set of moves
    \item $\lambda_A\from M_A\to \OP$ identifies each move as an $O$-move or a $P$-move
    \item $P_A\subset \seqs{M_A}$ is a set of legal positions
    \item $\zeta_A\from P_A\to\OP$ designates each position as an $O$-position or a $P$-position
  \end{itemize}
  \dots subject to a few rules
\end{frame}

\begin{frame}
  \frametitle{Advantages}
  \pause
  We can distinguish between computations that diverge and computations that converge, but may take an arbitrarily large number of steps to do so.

  \pause
  We do not have to check for infinite chattering explicitly when defining the divergences of a composition of strategies.

  \pause
  Composition is associative.
\end{frame}

\begin{frame}
  \frametitle{Fair PCF}
  \[
    \begin{array}{cc}
      \inferrule*{ }{(\lambda x.M)N\opto M[N/x]} & \inferrule*{ }{\Y_T M\opto M(\Y_T M)} \\
      \inferrule*{ }{\IfO \0 \opto \lambda x.\lambda y.x} & \inferrule*{ }{\IfO (\suc \n) \opto \lambda x.\lambda y.(y \n)} \\
      \inferrule*{M\opto M'}{\suc M\opto\suc M'} & \inferrule*{M\opto M'}{\IfO M\opto \IfO M'} \\
      \inferrule*{M \opto M'}{M N \opto M' N} & \inferrule*[right=$n\in\bN$]{ }{\wn \opto \n} \\
    \end{array}
    \]

\end{frame}

\begin{frame}
  \frametitle{Must-testing for fair PCF}
  If $M$ is a term of Fair PCF of ground type $\nat$, write $M\mustconverge$ if $M$ has no infinite evaluation paths $M\opto M_1\opto M_2\opto\cdots$.

  \pause
  Examples:
  \begin{itemize}
    \item $\Omega_\nat=\Y_\nat(\lambda x.x)\not\mustconverge$
      \pause
    \item $\IfO \wn \;\Omega_\nat \;\0\not\mustconverge$
      \pause
    \item $\Y_{\nat\to\nat} (\lambda f.\lambda x.\IfO x\;\0\;f)\wn\mustconverge$
  \end{itemize}
\end{frame}

\begin{frame}
  \pause
  \frametitle{Denotational Semantics for $\Y_T$}
  Harmer and McCusker define an order $\stle$ on strategies for a game $A$:

  $\sigma\stle\tau$ if
  \begin{enumerate}
    \item $\sigma\subset\tau$
    \item Every divergent position in $\tau$ is divergent in $\sigma$
    \item Every position in $\tau\setminus\sigma$ is divergent in $\sigma$.  
  \end{enumerate}

  \pause
  The order $\stle$ is preserved by composition.  

  \pause
  If $F$ is a set of strategies that is directed with respect to $\stle$ then $F$ has a least upper bound given by
  \[
    \left(\bigcup_{\sigma\in F}\sigma\;,\;\bigcap_{\sigma\in F}D_\sigma\right)
    \]
\end{frame}

\begin{frame}
  \frametitle{Denotational semantics for $\Y_T$}

  This means that we can define a strategy $\Y_A$ for $(A\iimpl A) \iimpl A$ to be the least fixed point of the strategy corresponding to the $\lambda$-term
  \[
    \lambda F.\lambda f.f(F f)
    \]
  \pause
  However, $\stle$-limits are not preserved by composition (in either direction), so it is hard to prove computational adequacy for this denotation of $\Y_T$.
\end{frame}

\section{Ordinal-Indexed Recursion}

\begin{frame}
  \frametitle{Introduction to Ordinal-Indexed Recursion}
  \pause
  In languages with no nondeterminism or finite nondeterminism, we can study $\Y_T$ by studying its finite approximants $\Y_T^n$, where:
  \[
    \Y_T^n M = \underbrace{M\cdots M}_{n}\Omega_T
    \]
  \pause
  If $Y_T M M_1\dots M_n\mustconverge$, then $\Y_T^n M M_1\dots M_n\mustconverge$ for some $n$ -- since branching is finite, any well-founded evaluation tree is \emph{bounded}.  

  \pause
  However, in the presence of infinite branching, we may have well-founded evaluation trees that are not bounded.  

  \pause
  For example, $\Y_{\nat\to\nat}(\lambda f.\lambda x.\IfO x\; \0\; f)\wn\mustconverge$, but $\Y_{\nat\to\nat}^n(\lambda f.\lambda x.\IfO x\; \0\; f)\wn\not\mustconverge$ for all $n$.  
\end{frame}

\begin{frame}
  \frametitle{Introduction to Ordinal-Indexed Recursion}
  We replace $\Y_T$ with a new family of constants $\Y_T^\alpha$, where $\alpha$ ranges over countable ordinals and a formal divergence symbol $\Omega_T$.  $\Y_T^\alpha$ has operational semantics given by:
  \[
    \inferrule*[right=$\beta<\alpha$]{ }{\Y_T^\alpha M\opto M(\Y_T^\beta M)}\hspace{24pt}
    \inferrule*{ }{\Y_T^0M\opto\Omega_{T}}
    \]
  \pause
  If $M\mustconverge$, then the evaluation tree for $M$ is well-founded and hence has complexity given by some countable ordinal $\alpha$.  If we replace $\Y$ with $\Y^\alpha$ in the description of $M$, then we can recover the original evaluation tree.  

  \pause
  However, the new term will have additional branches, some of which will diverge.  

  \pause
  We want the choice of ordinal $\beta$ to be `friendly', so that it does not affect the behaviour of $\mustconverge$.  
\end{frame}

\begin{frame}
  \frametitle{Big-Step semantics for $\mustconverge$ in Fair PCF}

  We can get round this problem using big step semantics.  

  \pause
  We add the following rule for $\wn$ to the usual big-step semantics for $\mustconverge$ in PCF:
  \[
    \inferrule{\forall n\in\omega\esuchthat M n\mustconverge}{M\wn\mustconverge}
    \]
  \pause
  We then give big-step semantics for the constants $\Y_T^\alpha$:
  \[
    \inferrule{\exists \beta<\alpha\esuchthat M(\Y_T^\beta M) M_1\cdots M_n\mustconverge}{\Y_T^\alpha MM_1\cdots M_n\mustconverge}
    \]
\end{frame}

\begin{frame}
  \frametitle{An example}
  \begin{IEEEeqnarray*}{Cl}
    &\Y_{\nat\to\nat}^\omega (\lambda f.\lambda x.\IfO x\;\0\;f)\;\wn\\
    \pause
    \opto &((\lambda f.\lambda x.\IfO x\;\0\;f) \Y_{\nat\to\nat}^n(\lambda f.\lambda x.\IfO x\;\0\;f))\;\wn\\
    \pause
    \opto &\IfO \wn\;\0\;(\Y^n_{\nat\to\nat}(\lambda f.\lambda x.\IfO x\;\0\;f)) \\
    \pause
    \opto &\IfO (n+1)\;\0\;(\Y^n_{\nat\to\nat}(\lambda f.\lambda x.\IfO x\;\0\;f))\\
    \pause
    \opto &\cdots\opto\Omega_{\nat}
    \pause
  \end{IEEEeqnarray*}
  \begin{IEEEeqnarray*}{Cl}
    &\Y_{\nat\to\nat}^{\omega+1}(\lambda f.\lambda x.\IfO x\;\0\;f)\;\wn\\
    \pause
    \opto &((\lambda f.\lambda x.\IfO x\;\0\;f) \Y_{\nat\to\nat}^\omega(\lambda f.\lambda x.\IfO x\;\0\;f))\;\wn\\
    \pause
    \opto &\IfO \wn\;\0\;(\Y^\omega_{\nat\to\nat}(\lambda f.\lambda x.\IfO x\;\0\;f)) \\
    \pause
    \opto &\IfO m\;\0\;(\Y^\omega_{\nat\to\nat}(\lambda f.\lambda x.\IfO x\;\0\;f))\\
    \pause
    \opto &\Y^\omega_{\nat\to\nat}(\lambda f.\lambda x.\IfO x\;\0\;f)\;m\\
    \pause
    \opto &(\lambda f.\lambda x.\IfO x\;\0\;f)\Y_{\nat\to\nat}^{m+1}(\lambda f.\lambda x.\IfO x\;\0\;f)\;m\\
    \pause
    \opto&\cdots\opto\0
  \end{IEEEeqnarray*}
\end{frame}

\begin{frame}
  \frametitle{Can we model $\Y_T^\alpha$ within the language?}
  \pause
  Recall that we could model $\Y_T^n M$ within PCF as
  \[
    \underbrace{M\cdots M}_{n}\Omega_T
    \]
  \pause
  We can do this uniformly if we add a recursion constant $\rec_T\from \nat\to T\to(T\to T)\to T$ to the language.  Then we can model
  \[
    \Y_T^n = \lambda F.\rec_T n\;\Omega_T\;F
    \]
  \pause
  But how would we model $\Y_T^\omega$?  We need to add friendly choice to the language....
\end{frame}

\begin{frame}
  \frametitle{Explicit friendly choice}
  \pause
  We add a new constant to our language:
  \[
    \nw\from\nat
    \]
  \pause
  We give big-step semantics for $\mustconverge$ corresponding to $\nw$:
  \[
    \inferrule{\exists n\in\omega\esuchthat Mn\mustconverge}{M\nw\mustconverge}
    \]
  Then we may model $\Y_T^\omega$ as:
  \[
    \Y_T^\omega = \lambda F.\rec_T \nw\;\Omega_T\;F
    \]
  \pause
  Going further, we may model $\Y_T^{\omega+1}$ as $\lambda F.F(\rec_T\nw\;\Omega_T\;F)$, and so on.  

  \pause
  How far can we go?
\end{frame}

\begin{frame}
  \frametitle{Defining recursion principles using friendly choice}
  \pause
  Suppose that we can define an ordinal $\alpha$ in our language as a Church encoding:
  \[
    \alpha_T\from T \to (T \to T) \to ((\nat \to T) \to T) \to T
    \]
  \pause
  Then we can simulate $\Y_T^\alpha$ within our language:
  \[
    \Y_T^\alpha\equiv \lambda F.\alpha_T\;\Omega_T\;F\;(\lambda f.f\nw)
    \]
  \pause
  Using  $\rec_T$, it is possible to define Church encodings $\alpha_T$ for all ordinals $\alpha<\epsilon_0$, so we can simulate $\Y_T^\alpha$ for all $\alpha<\epsilon_0$ in PFC - $\Y$ + $\rec$ + $\nw$.
\end{frame}

\section{Game Semantics for Friendly Choice}

\begin{frame}
  \frametitle{Friendly vs. unfriendly choice: a 2-player game within the language}
  \pause
  Suppose we have a constant
  \begin{gather*}
    \Iflt\from\nat\to\nat\to\nat\to\nat\to\nat
  \end{gather*}
  in our language with the obvious meaning.

  \pause
  Consider the terms
  \begin{gather*}
    \Iflt \wn\;\nw\;\0\;\Omega_\nat\\
    \Iflt \nw\;\wn\;\Omega_\nat\;\0
  \end{gather*}
  \pause
  Because of evaluation order, the first term must converge, but the second may diverge.

  \pause
  With more complicated predicates, we can encode more complicated (bounded) Gale-Stewart games in the language.
\end{frame}

\begin{frame}
  \frametitle{Game Semantics for Friendly Choice}
  \pause
  We need the ability to make all choices at once, but forget about the bad ones (unless all our choices were bad).  

  \pause
  If $A$ is a game, then a \emph{flexible strategy} for $A$ is a pair $\Sigma=(\sigma_\Sigma,F_\Sigma)$, where $\sigma_\Sigma$ is a strategy for $A$ and $F_\Sigma\subset\sigma_\Sigma$ is a set of positions in $\sigma_\Sigma$ that are forced.  All other positions in $\sigma$ are optional.

  \pause
  Then a \emph{switching} for $\Sigma$ is a strategy $\sigma'$ for $A$ such that $\sigma'\subset\sigma_\Sigma$ and such that if $s\in\sigma'$ and $sab\in F_{\Sigma}$ then $sab\in\sigma'$.  In this case, we write $\sigma'\sw\Sigma$.

  \pause
  Given flexible strategies $\Sigma\from A\to B$ and $\Tau\from B\to C$, we may form the composite:
  \[
    \comp{\Tau}{\Sigma} = \exlub_{\substack{\sigma'\sw\Sigma\\\tau'\sw\Tau}}\comp{\tau'}{\sigma'}
    \]
\end{frame}

\begin{frame}
  \frametitle{Game semantics for friendly choice}
  \pause
  We embed the original game semantics into the flexible game semantics by sending $\sigma\mapsto|\sigma|\coloneqq(\sigma,\sigma)$.  

  \pause
  So $\wn$ inherits its original semantics as $(\wn, \wn)$.  

  \pause
  We give $\nw$ an explicitly flexible semantics:
  \[
    \deno{\nw} = (\wn, \{\emptyplay\})
    \]
  \pause
  So all the choices of number in $\nw$ are optional.  
  \pause
  We get a consistency result:
  \begin{proposition}
    If $M$ is a term of PCF - $\Y$ + $\wn$ + $\nw$ of type $\nat$ such that $M\mustconverge$, then there are no divergent positions in $\deno{M}$.  
  \end{proposition}
\end{frame}

\begin{frame}
  \frametitle{Stable ordering for flexible strategies}
  \pause
  We can extend the stable order $\stle$ to flexible strategies.

  \pause
  We say that $\Sigma\stle\Tau$ if $\sigma_\Sigma\stle\sigma_\Tau$ and $\sigma_\Sigma\cap F_\Tau\subset F_\Sigma$.  

  \pause
  Then we have least upper bounds for stably-directed sets $X$ of flexible strategies:
  \begin{gather*}
    \sigma_{\exlub_{\Sigma\in X}\Sigma} = \exlub_{\Sigma\in X}\sigma_\Sigma \\
    F_{\exlub_{\Sigma\in X}\Sigma} = \bigcap_{\Sigma\in X}F_\Sigma
  \end{gather*}
  \pause
  $\exlub_{\Sigma\in F}\Sigma$ is a `friendly choice between the $\Sigma\in F$'.  

  \pause
  We can then construct the $\Y_T^\alpha$ as the elements of the Bourbaki-Witt chain.  

  \pause
  For suitably large $\alpha$ (conjecturally $\alpha=\epsilon_0$), we get a sound and adequate semantics for $\Y_T$.  
\end{frame}

\end{document}
