\documentclass{beamer}

\let\BEAMER=1
\let\NOTARTICLE=1
\let\FEWFONTS=1

\input{../../common/preamble}

\renewcommand{\read}{{\mathsf{read}}}
\newcommand{\wwrite}{{\mathsf{write}}}
\newcommand{\Read}{{\mathsf{Read}}}
\newcommand{\Write}{{\mathsf{Write}}}
\newcommand{\tick}{\checkmark}
\renewcommand{\int}{\texttt{int}}

\title{A Coalgebraic Approach to Stateful Objects in Game Semantics}

\author{John Gowers and James Laird}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}
  \frametitle{Semantics of Stateful Programs}
  In a \emph{purely functional} language, we can represent a function of type $S\to T$ as a function between sets:
  \[
    \deno{f}\from\deno{S}\to\deno{T}
    \]
  \pause
  In a stateful language, we have to take the program \emph{state} into consideration.
  \pause
  We represent a stateful function as a function:
  \[
    \deno{f}\from W \times \deno{S}\to W \times \deno{T}
    \]
  where $W$ denotes the program state.
\end{frame}

\begin{frame}
  \frametitle{Game Semantics}
  In Game Semantics, we denote types by games and we denote elements of types by strategies for those games.  
  \pause
  We start with very simple games and build up more complicated ones using \emph{connectives}
  \pause
  For example, if $X$ is a set of values, we have a game $\Read[X]$ and a game $\Write[X]$:
  \[
    \begin{array}{cc}
      \Read[X]:\ \wn x\ (x\in X)
        & \Write[X]:\  x\tick\ (x\in X)
    \end{array}
    \]
  \pause
  Given games $A,B$, we can form the \emph{compound games} $A\times B$, $A\tensor B$, $A\implies B$, $A\sequoid B$,... in which the games $A$ and $B$ are played in parallel.
\end{frame}

\begin{frame}
  \frametitle{Function types and Exponentials}
  Games of the form $A\implies B$ represent functions from $A$ to $B$.
  \pause
  \[
    \begin{array}{ccc}
      \Read[\bN] & \implies & \Read[\bN] \\
      \pause
      & & \wn \\
      \pause
      \wn & & \\
      \pause
      7 & & \\
      \pause
      & & 49
    \end{array}
    \]
  \pause
  But what if the argument is a function?
  \pause
  For example, take the lambda term
  \[
    \lambda f^{\nat\to\nat}.f(f\ 3)\from (\nat\to\nat)\to\nat
    \]
\end{frame}

\begin{frame}
  \frametitle{Function types and Exponentials}
  \[
    \lambda f^{\nat\to\nat}.f(f\ 3)\from (\nat\to\nat)\to\nat
    \]
  \pause
  \[
    \begin{array}{ccccc}
      (\Read[\bN] & \implies & \Read[\bN]) & \implies & \Read[\bN] \\
      \pause
      & & & & \wn \\
      \pause
      & & \wn & & \\
      \pause
      \wn & & & & \\
      \pause
      3 & & & & \\
      \pause
      & & 9 & & \\
    \end{array}
    \]
  \pause
  ...but now we are stuck.
  \pause
  We introduce the \emph{exponential} $\oc A \approx A \sequoid (A \sequoid (A \sequoid \cdots$
  \pause
  Now we represent the type of functions from $A$ to $B$ using the game
  \[
    \oc A \implies B
    \]
\end{frame}

\begin{frame}[fragile]
  \frametitle{A storage cell}
  Consider the following (Java) class:
  \begin{verbatim}
  public class StorageCell implements AbstractCell {
    private int state;
    public int read() { return state; }
    public void write(int x) { state = x; }
  }
  \end{verbatim}
  \pause
  We represent the \emph{public interface} of the class with the game
  \[
    \oc(\Write[\int] \times \Read[\int])
    \]
  \pause
  The \emph{implementation} now corresponds to a \emph{strategy} for that game:
  \[
    7\ \tick\ \wn\ 7\ \wn\ 7\ 8\ \tick\ 9\ \tick\ \wn\ 9\dots
    \]
\end{frame}

\begin{frame}[fragile]
  \frametitle{A state-transformer based approach}
  \pause
  The strategy is easy to define, but it is hard to study because the state is \emph{implicit} or \emph{encapsulated}.  
  \pause
  We want to return to the classical state-transformer approach
  \pause
  \tikzset{dotsstyle/.style={draw=none}}
  \[
    \begin{tikzpicture}
      \matrix[matrix of nodes, column sep=36pt, inner xsep=1ex, name=m, nodes={draw=black}] {
        & in & in & in & & \\
        $A_0$ & $A_1$ & $A_2$ & $A_3$ & |[dotsstyle]| $\cdots$ \\
        & out & out & out & & \\};
      \path[commutative diagrams/.cd, every arrow, every label, labels={font size=tiny}]
        (m-2-1) edge node {init} (m-2-2)
        (m-2-2) edge node {state} (m-2-3)
        (m-2-3) edge node {state} (m-2-4)
        (m-2-4) edge node {state} (m-2-5);
    \end{tikzpicture}
    \]
  \pause
  We may construct a state transformer (strategy) $\oc X \implies (\Read[X] \times \Write[X]) \sequoid \oc X$ corresponding to a single invocation of the Java class on the previous slide.
  \pause

  Using some category-theoretic magic ($\oc A$ is the \emph{final coalgebra} for the endofunctor $A\sequoid\blank$), we can automatically construct our strategy for $\oc X \implies \oc (\Read[X]\times \Write[X])$ from this state transformer.
\end{frame}

\end{document}
